---
title: "Introduction to MetaAnalysis"
author: "TODO"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Introduction
------------

The purpose of this package is to take a series of data sets containing p biomarkers as represented in the columns of the data frame, and calculate a series of p-values for each column, as well as run a pooling method based on user input.

Installing
----------

The easiest way to get mega.analysis is to install the whole package, use the following :

`install.packages(mega.analysis)`

And then to load it, use the following:

`library(mega.analysis)`

Workflow
--------

The package will not take less than two data frames and no more than five. In addition, the package will not run unless the number of columns between all the data frames is the same. The amount rows in each data set do not have to be the same, however, each row must be organized into a numeric group in a column placed before all the biomarkers. The number of groups in each data frame must also all be the same. 

The package will not take less than two data frames and no more than five. In addition, the package will not run unless the number of columns between all the data frames is the same. The amount rows in each data set do not have to be the same, however, each row must be organized into a numeric group in a column placed before all the biomarkers. The number of groups in each data frame must also all be the same. 

The second function of this package is to take the collection of p-values and pool them by biomarker. Each biomarker will undergo pooling based on four tests (As highlighted in the “Pooling Methods” Section), the Fisher, Stouffer, Minimum p-value (minP), and Maximum p-value (maxP). If one of these tests is specified in the “type” parameter, the package will then provide a list of p-values based on that test. If any other input is made, the package will output all the tests at once.

you can use: `meta.analysis(...,method)` to compute pooled p value for different biomarkers by using four different methods: Fisher, Stouffer, Maxium p value and Minium P value.

**Pooling Commands**

+ Fisher: set 	method=”Fisher”
+ Stouffer: set	 method=”Stouffer”
+ MinP: set 	method=”minP”
+ MaxP: set 	method=”maxP”

```{r, eval=F}
set.seed(123)
p <- 10
data1 <- data.frame(group=sample(1:3,20,replace=TRUE),matrix(rnorm(p*20),ncol=p))
data2 <- data.frame(group=sample(1:2,15,replace=TRUE),matrix(rnorm(p*15),ncol=p))
meta.analysis(data1,data2)
```


*Please note that setting the type is case sensitive, and will default to all cases if not. Ex: typing “stouffer” instead of “Stouffer” will output all tests instead of just the Stouffer test.*

# Pooling Methods

1. Fisher's Method
  The Fisher's method sums up the log-transformed p-values obtained from individual studies.Fisher's statistic $\chi_{Fisher}^{2} = -2 \sum_{i=1}^K log(p_i)$ follows a $\chi^{2}$ distribution with $2K$ degrees of freedom under the null hypothesis. Smaller p-values contribute larger scores to the Fisher's statistic.
  
2. Stouffer's Method
  Stouffer's methods sums the inverse normal trasformed p-values. Stouffer's statistic $T\_Stouffer = \sum_{i=1}^K \frac{\Phi^{-1}(p_i)}{\sqrt{K}}$ ( $\Phi$ is the cdf of a standard normal distribution) follows a standard normal distribution under the null hypothesis. Similar to Fisher's method, smaller p-values contribute more to the Stouffer's score.
  
3. Minimum p-value Method
  The minimun p-value method take the minimum p-value among the K studies as the test statistic. It follows a beta distribution with degrees of freedom $\alpha = 1$ and $\beta = K$ under the null hypothesis.
  
4. Maximum p-value Method
  The maximum p-value method take the maximum p-value as the test statistic. It follows a beta distribution with degrees of freedom $\alpha = K$ and $\beta = 1$ under the null hypothesis.
 
# Example

1. Input dataframes
  * The number of input dataframes should between 2 and 5
  * Each dataframe should have the same number of columns
  * The function mata.analysis() will automatically check if all inputs meet requirements
```{r}
library(MetaAnalysis)
set.seed(123)
p <- 20
# dataframe1 has 20 columns and 3 groups
data1 <- data.frame(group=sample(1:3,300,replace=TRUE),matrix(rnorm(p*300),ncol=p))
# dataframe1 has 20 columns and 2 groups
data2 <- data.frame(group=sample(1:2,200,replace=TRUE),matrix(rnorm(p*200),ncol=p))
# dataframe1 has 20 columns and 5 groups
data3 <- data.frame(group=sample(1:5,500,replace=TRUE),matrix(rnorm(p*500),ncol=p))
```
  * Input data error: if number of dataframes are out of limit, or number of columns are not equal.

2. Obtaining p-value matrix
  * #######If not specified, the default level of confidence is $\alpha = 0.05$ ######
```{r, eval=FALSE}
meta.analysis(data1, data2, data3, method='test', alpha = 0.01)$p.matrix
```
3. Pooling p-values
  * 4 pooling methods introduced above are available
  * The default results will show Fisher, Stouffer, minP, and maxP in a list
```{r, eval=FALSE}
meta.analysis(data1, data2, data3, method='test')$pooled.p.matrix
```
4. Tests performed
  * User can get a record of the results from tests performed. For example, shapiro test for normality. 
```{r, eval=FALSE}
meta.analysis(data1, data2, data3, method='test')$test.performed
```
  
