---
title: "Introduction to MetaAnalysis"
author: "Yen-An Chen, Wendy Lian, Siwei Chen, Peter Gonatas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The purpose of this package is to take a series of data sets containing $p$ biomarkers as represented in the columns of the data frame, and calculate a series of p-values for each column, as well as run a pooling method based on user input.

# Installing

To install MetaAnalysis package from source, open terminal application and enter command:

```bash 
R CMD INSTALL [PATH_TO_FILE]/MetaAnalysis_0.1.tar.gz
```

# Workflow

The package will not take less than two data frames and no more than five. In addition, the package will not run unless the number of columns between all the data frames is the same. The amount rows in each data set do not have to be the same, however, each row must be organized into a group in a column placed before all the biomarkers. 

The second function of this package is to take the collection of p-values and pool them by biomarker. Each biomarker will undergo pooling based on four tests (As highlighted in the “Pooling Methods” Section), the Fisher, Stouffer, Minimum p-value (minP), and Maximum p-value (maxP). If one of some of these tests is specified in the “method” parameter, the package will then provide a matrix of p-values based on these tests.

you can use: `meta.analysis(...,method)` to compute pooled p value for different biomarkers by using four different methods: Fisher, Stouffer, Maxium p value and Minium P value.

`method` can be a single string or a vector of the collection of the following options:

**Pooling Commands**

* Fisher: set method=`'Fisher'`
* Stouffer: set method=`'Stouffer'`
* MinP: set	method=`'minP'`
* MaxP: set	method=`'maxP'`

*Please note that setting the type is case sensitive, and will default to all methods, i.e. `c('Fisher', 'Stouffer', 'minP', 'maxP')`*

# Pooling Methods

1. Fisher's Method
  The Fisher's method sums up the log-transformed p-values obtained from individual studies. Fisher's statistic $\chi_{Fisher}^{2} = -2 \sum_{i=1}^K \log(p_i)$ follows a $\chi^{2}$ distribution with $2K$ degrees of freedom under the null hypothesis. Smaller p-values contribute larger scores to the Fisher's statistic.
  
2. Stouffer's Method
  Stouffer's methods sums the inverse normal trasformed p-values. Stouffer's statistic $T = \sum_{i=1}^K \frac{\Phi^{-1}(p_i)}{\sqrt{K}}$ ( $\Phi$ is the cdf of a standard normal distribution) follows a standard normal distribution under the null hypothesis. Similar to Fisher's method, smaller p-values contribute more to the Stouffer's score.
  
3. Minimum p-value Method
  The minimun p-value method take the minimum p-value among the $K$ studies as the test statistic. It follows a beta distribution with degrees of freedom $\alpha = 1$ and $\beta = K$ under the null hypothesis.
  
4. Maximum p-value Method
  The maximum p-value method take the maximum p-value as the test statistic. It follows a beta distribution with degrees of freedom $\alpha = K$ and $\beta = 1$ under the null hypothesis.
 
# Example

1. Input dataframes
    + The number of input dataframes should between 2 and 5
    + Each dataframe should have the same number of columns
    + The function `mata.analysis()` will automatically check if all inputs meet requirements
    + If not specified, the default level of confidence is $\alpha = 0.05$

```{r}
library(MetaAnalysis)
set.seed(123)  ## Fix seed for reproducible result
p <- 5
# dataframe1 has 5 columns and 3 groups
data1 <- data.frame(group=sample(1:3,200,replace=TRUE),matrix(rnorm(p*200),ncol=p))
head(data1)

# dataframe1 has 5 columns and 2 groups
data2 <- data.frame(group=sample(1:2,150,replace=TRUE),matrix(rnorm(p*150),ncol=p))
# dataframe1 has 5 columns and 4 groups
data3 <- data.frame(group=sample(1:4,400,replace=TRUE),matrix(rnorm(p*400),ncol=p))


res = meta.analysis(data1, data2, data3, method=c('Fisher', 'Stouffer', 'minP', 'maxP'), alpha = 0.05)
```
*Input data error: if number of dataframes are out of limit, or number of columns are not equal.*

2. Obtaining p-value matrix
```{r}
res$p.matrix
```

3. Tests performed
    + User can get a record of the results from tests performed.
```{r}
res$test.performed
```
  
4. Pooling p-values
    + 4 pooling methods introduced above are available.
    + The default results will show Fisher, Stouffer, minP, and maxP in a matrix.
```{r}
res$pooled.p.matrix
```

